---
title: "Add geocodes"
output: html_notebook
---

Data from Chicago, Detroit, Los Angeles, NYC, San Francisco, Seattle, Tucson and 
Virginia Beach are already geocoded. Co-ordinates for the remaining cities are
added here, usually based on data from [Geocodio](http://geocod.io).


# Fort Worth

Many offences in Fort Worth are not geocoded. It appears that some locations are 
sometimes geocoded but sometimes not. In fact, most locations that are not 
geocoded are geocoded elsewhere in the dataset. However, looking at a list of 
locations shows that some (e.g. '100 BLOCK CALHOUN ST') have multiple locations 
that are too far from one another to be the product of geocoding issues. E.g. in 
the case of Calhoun St, the co-ordinates suggest some offences actually occurred 
at South Calhoun St. This means we can't apply geocodes from a location to 
offences at the 'same' address but which have no geocodes. It will therefore be 
necessary to geocode all the missing locations.

```{r}
# read crime data
data_fw <- readRDS("../temp_data/raw_fort_worth_data.Rds")
```

## Export un-geocoded offences

```{r}
# attempt to make sense of the address abbreviations used in FW
data_fw$`Block Address Expanded` <- data_fw$`Block Address` %>% 
  str_to_upper() %>% 
  str_replace_all('\\s+', ' ') %>%
  str_replace_all('\\bBLOCK [DGHKPRVY]\\b', '') %>%
  str_replace_all('\\b(BLOCK|RA|RAMP|NB|EB|SB|WB)\\b', '') %>%
  str_replace_all('\\bFWY\\b', 'FREEWAY') %>%
  str_replace_all('\\bIH\\b', 'INTERSTATE') %>%
  str_replace_all('\\bSR\\b', 'SERVICE ROAD') %>%
  str_replace_all("\\bWA\\b", "WAY") %>% 
  str_replace_all('\\bBELL SP\\b', 'BELL HELICOPTER BL') %>%
  str_replace_all('\\bLOOP 830\\b', 'LOOP INTERSTATE 820') %>%
  str_replace_all(
    "\\b(NORTH|EAST|SOUTH|WEST|NORTHEAST|NORTHWEST|SOUTHEAST|SOUTHWEST) LOOP\\b", 
    "\\1 LOOP INTERSTATE 820"
  ) %>%
  str_replace_all('\\d+$', '') %>%
  str_replace_all('\\s+', ' ') %>%
  str_trim()

# export unique addresses that do not have geocodes to CSV for geocoding
data_fw %>%
  filter(is.na(Latitude) | is.na(Longitude)) %>%
  group_by(`Block Address`, `Block Address Expanded`, `City`, `State`) %>% 
  summarise(n = n()) %>%
  write_csv('../geocoding_data/fw_missing_locations.csv')
```

## Merge co-ordinate data

```{r}
# import geocoded data from CSV files (there are multiple files because of usage 
# limits on the geocoding service)
locations <- dir("../geocoding_data", pattern = 'fw_geocoded_locations*', 
                 full.names = TRUE) %>% 
  map(read_csv, col_types = cols(.default = col_character())) %>%
  bind_rows() %>% 
  # filter out locations that could only be geocoded to a city rather than to a
  # street or point
  filter(!`Accuracy Type` %in% c('place', 'state')) %>% 
  distinct(street, .keep_all = TRUE) %>% 
  select(`Block Address Expanded` = street, NewLatitude = Latitude,
         NewLongitude = Longitude)

# merge datasets and add co-ordinates where necessary
data_fw <- data_fw %>% 
  left_join(locations, by = 'Block Address Expanded') %>% 
  mutate(
    Longitude = ifelse(is.na(Longitude), NewLongitude, Longitude) %>% 
      as.double(),
    Latitude = ifelse(is.na(Latitude), NewLatitude, Latitude) %>% as.double()
  ) %>% 
  # Remove fields created during this procedure, to avoid cluttering the final 
  # dataset
  select(-NewLatitude, -NewLongitude) 

# identify failed geocoding
data_fw %>% 
  filter(is.na(Latitude) | is.na(Longitude)) %>%
  group_by(`Block Address Expanded`, Date.Year) %>% 
  summarise(n = n()) %>% 
  spread(Date.Year, n) %>% 
  ungroup() %>% 
  mutate(Total = rowSums(select_if(., is.numeric), na.rm = TRUE)) %>% 
  arrange(desc(Total)) %>% 
  write_csv("../geocoding_data/fw_geocoding_failures.csv", na = "")

# report success
data_fw %>% 
  mutate(geocoded = ifelse(!is.na(Latitude) & !is.na(Longitude), TRUE, 
                           FALSE)) %>% 
  group_by(geocoded) %>% 
  summarise(n = n()) %>% 
  { cat(format(.$n[.$geocoded == FALSE], big.mark = ","), "cases –", 
        format((.$n[.$geocoded == FALSE] / sum(.$n)) * 100, digits = 2), 
        "% of the total – could not be geocoded\n") }

# save crime data
saveRDS(data_fw, "../temp_data/raw_fort_worth_data.Rds")

# clean up
rm(locations, data_fw)
```


# Louisville

No Louisville offences are geocoded, so it will all have to be done manually.

```{r}
data_lo <- readRDS("../temp_data/raw_louisville_data.Rds")
```


```{r}
# remove quirks that might make geo-coding more difficult
data_lo$BLOCK_ADDRESS <- data_lo$BLOCK_ADDRESS %>% 
  str_to_upper() %>% 
  # remove 'BLOCK' from each address
  str_replace_all('\\bBLOCK\\b', '') %>% 
  # offences at a particular stadium
  str_replace_all('^@WF -\\b', 'WATERFRONT PARK') %>% 
  # offences in parks
  str_replace_all('^@PARK -\\b', '') %>%
  # offences at interstate on/off ramps
  str_replace_all('^@(\\d+\\.*\\d*)\\b(.*)$', '\\2 EXIT \\1') %>% 
  # other offences starting @, mostly in the form of cross-streets
  str_replace_all('^@', '') %>% 
  # offences on the zero block of a street
  str_replace_all('^0\\b', '1') %>% 
  str_replace_all('\\b(AT|TO)\\b', ' / ') %>% 
  str_replace_all('\\s+', ' ') %>%
  str_trim()

# add city and state variables
data_lo$geocode_city <- 'Louisville'
data_lo$geocode_state <- 'Kentucky'

# export random sample of data for testing
data_lo %>%
  filter(BLOCK_ADDRESS != 'COMMUNITY AT LARGE') %>%
  group_by(BLOCK_ADDRESS, geocode_city, geocode_state, ZIP_CODE) %>% 
  slice(sample(1:n(), 100, replace = TRUE)) %>%
  write_csv('../geocoding_data/louisville_missing_locations_sample.csv')

# export unique addresses that do not have geocodes to CSV for geocoding
data_lo %>%
  filter(BLOCK_ADDRESS != 'COMMUNITY AT LARGE') %>%
  group_by(BLOCK_ADDRESS, geocode_city, geocode_state, ZIP_CODE) %>% 
  summarise(n = n()) %>%
  write_csv('../geocoding_data/louisville_missing_locations.csv')

# remove variables added for this purpose
data_lo$geocode_city <- NULL
data_lo$geocode_state <- NULL
```

```{r}
locations <- read_csv('../geocoding_data/louisville_geocoded_locations.csv', 
                      col_types = cols(.default = col_character())) %>% 
  filter(!`Accuracy Type` %in% c('place', 'state') & `State` == 'KY') %>% 
  distinct(BLOCK_ADDRESS, .keep_all = TRUE) %>% 
  select(BLOCK_ADDRESS, Latitude, Longitude)

data_lo <- data_lo %>% 
  left_join(locations, by = 'BLOCK_ADDRESS') %>% 
  mutate(
    Latitude = as.double(Latitude),
    Longitude = as.double(Longitude)
  )

# identify failed geocoding
data_lo %>% 
  filter(is.na(Latitude) | is.na(Longitude)) %>%
  group_by(BLOCK_ADDRESS, Date.Year) %>% 
  summarise(n = n()) %>% 
  spread(Date.Year, n) %>% 
  ungroup() %>% 
  mutate(Total = rowSums(select_if(., is.numeric), na.rm = TRUE)) %>% 
  arrange(desc(Total)) %>% 
  write_csv("../geocoding_data/louisville_geocoding_failures.csv", na = "")

# report success
data_lo %>% 
  mutate(geocoded = ifelse(!is.na(Latitude) & !is.na(Longitude), TRUE, 
                           FALSE)) %>% 
  group_by(geocoded) %>% 
  summarise(n = n()) %>% 
  {
    cat(format(.$n[.$geocoded == FALSE], big.mark = ","), "cases –",
        format((.$n[.$geocoded == FALSE] / sum(.$n)) * 100, digits = 2), 
        "% of the total – could not be geocoded\n")
  }

# save crime data
saveRDS(data_lo, "../temp_data/raw_louisville_data.Rds")

# clean up
rm(locations, data_lo)
```

