---
title: "Spatial processing"
output: html_notebook
---

This file takes the geocoded processed data for each city and carries out 
spatial processing.


# Convert all co-ordinates to lat/lon pairs

Data for most cities already include lat/lon pairs, but some require conversion
from other co-ordinate systems.

## Detroit

Detroit data have lat/lon pairs, but they seem to encode NA as 999999.0 degrees. 
These cases can easily be rectified by replacing any value of latitude that is 
not between -90 and +90, and any value of longitude that is not between -180 and 
+180, with NA.

```{r}
data_dt <- readRDS("../temp_data/raw_detroit_data.Rds")

data_dt <- data_dt %>% 
  mutate(
    Latitude = ifelse(Latitude < -90 | Latitude > 90, NA, Latitude),
    Longitude = ifelse(Longitude < -180 | Longitude > 180, NA, Longitude)
  )
  
saveRDS(data_dt, "../temp_data/raw_detroit_data.Rds")

rm(data_dt)
```


# Stip offences outside cities and add census identifiers

```{r}
setdiff(
  dir("../temp_data", pattern = "^raw_") %>% 
    str_extract("\\_([a-z_]+?)\\_data.Rds") %>% 
    as.character(),
  dir("../temp_data", pattern = "^spatial_") %>% 
    str_extract("\\_([a-z_]+?)\\_data.Rds") %>% 
    as.character()
) %>% 
  { paste0("../temp_data/raw", .) } %>% 
  sapply(function (x) {
  
  # get city name
  city_name <- str_match(x, "raw\\_([a-z_]+?)\\_data.Rds") %>% 
    as.character() %>% 
    last() %>% 
    str_replace_all("\\_", " ") %>% 
    str_to_title()
  
  start_time <- now()
  cat("\nProcessing", city_name, "data\nStarting at", 
      format(start_time, "%F %T"), "\n")
  
  # get city metadata
  city_fips <- cities$fips[cities$name == city_name]
  
  # load data
  data <- readRDS(x)%>% 
    mutate(
      Latitude = as.numeric(Latitude),
      Longitude = as.numeric(Longitude)
    )
  
  # note number of rows in data
  initial_rows <- nrow(data)
  cat("✔︎ Loaded data with", format(initial_rows, big.mark = ","),
      "rows\n")
  
  # convert to SF object (removing any rows without co-ordinates)
  offences <- data %>% 
    select(UID, Longitude, Latitude)  %>% 
    filter(!is.na(Longitude) & !is.na(Latitude)) %>% 
    st_as_sf(coords = c('Longitude', 'Latitude')) %>% 
    st_set_crs(4326)
  no_coords <- initial_rows - nrow(offences)
  if (nrow(offences) > initial_rows) {
    stop('✖ Rows have been added to data', call. = FALSE)
  } else {
    cat("✔︎ Converted data to an SF object, removing", 
        format(no_coords, big.mark = ","), 
        "rows without co-ordinates\n")
  }
  
  # create a spatial object for the city outline
  outline <- st_read(
    paste0("../spatial_data/cities/tl_2016_", city_fips, "_place.shp"), 
    quiet = TRUE, stringsAsFactors = FALSE) %>% 
    filter(NAME == city_name) %>%
    st_transform(4326)
  cat("✔︎ Created SF object for city outline\n")
  
  # identify whether each offence is within the city boundary
  cat("🕣 Identifying rows outside the city\n")
  offences$in_city <- st_covered_by(offences, outline, sparse = FALSE) %>% 
    as.logical()
  
  # filter out offences outside the city boundary
  outside_city <- nrow(offences) - sum(offences$in_city) - no_coords
  # offences <- offences %>% filter(in_city == TRUE)
  
  if (nrow(offences) > initial_rows) {
    stop('✖ Rows have been added to data', call. = FALSE)
  } else {
    cat("✔ Identified", format(outside_city, big.mark = ","), 
        "rows that were outside the city\n")
  }

  # create a spatial object for census blocks
  blocks <- st_read(
    paste0("../spatial_data/cities/tl_2016_", city_fips, "_tabblock10.shp"), 
    quiet = TRUE, stringsAsFactors = FALSE) %>%
    select(
      fips_state = STATEFP10, # FIPS state code
      fips_county = COUNTYFP10, # FIPS county code
      Tract = TRACTCE10, # Census 2010 tract ID
      Block = BLOCKCE10 # Census 2010 block ID
    ) %>% st_transform(4326)
  cat("✔︎ Created SF object for census blocks\n")
  
  # join census blocks to offences
  # Some crimes will fall on the boundary of multiple census blocks, so the
  # following code simply takes the first block in each case
  cat("🕣 Identifying census blocks\n")
  offences <- offences %>% 
    st_join(blocks) %>% 
    mutate(block_code = paste0(fips_state, fips_county, Tract, Block)) %>%
    group_by(UID) %>% 
    top_n(1, block_code) %>% 
    ungroup() %>% 
    select(-block_code)

  if (nrow(offences) > initial_rows) {
    stop('✖ Rows have been added to data', call. = FALSE)
  } else {
    cat("✔ ︎Identified census blocks for", 
        format(nrow(offences), big.mark = ","), "rows\n")
  }

  # identify block groups
  offences <- offences %>% 
    mutate(block_group = str_sub(Block, 1, 1))
  cat("✔ ︎Identified census block groups for", 
      format(nrow(offences), big.mark = ","), "rows\n")

  # add the fields from the spatial joins to the existing crime data
  cat("🕣 Joining census identifiers to crime data\n")
  data <- data %>% left_join(offences, by = 'UID')
  
  if (nrow(data) > initial_rows) {
    stop('✖ Rows have been added to data', call. = FALSE)
  } else {
    cat("✔ Joined census identifiers to", 
        format(nrow(data), big.mark = ","), "rows of crime data\n")
  }
  
  # store data
  saveRDS(data, paste0("../temp_data/spatial_", 
          str_replace_all(str_to_lower(city_name), "\\s", "_"), "_data.Rds"))
  cat("✔ Saved data to file\n")
  
  duration <- difftime(now(), start_time, units = "auto")
  cat("Finished at", format(now(), "%F %T"), "after", 
      round(duration, digits = 1), units(duration), "\n")

}) %>% invisible()
```

# Check data

```{r}
for (city in cities$name[2]) {
  
  cat("\nProducing map for", city, "at", format(now(), "%F %T"), "\n")
  
  crimes <- readRDS(
    paste0("../temp_data/spatial_", 
           str_replace_all(str_to_lower(city), "\\s", "_"), "_data.Rds")
  ) %>% 
    sample_frac(0.1) %>% 
    # sample_n(1000) %>%
    filter(!is.na(Latitude) & !is.na(Longitude)) %>% 
    st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326)
  outline <- st_read(
    paste0("../spatial_data/cities/tl_2016_", cities$fips[cities$name == city], 
           "_place.shp"), 
    quiet = TRUE, stringsAsFactors = FALSE) %>% 
    filter(NAME == city)
  cat("\tSampled", format(nrow(crimes), big.mark = ","), 
      "rows (10% of total)\n")
  outline_bbox <- st_bbox(outline)
  
  ggplot() +
    geom_sf(data = outline, fill = NA) +
    geom_sf(aes(shape = in_city), data = crimes, alpha = 0.5) +
    theme_minimal() +
    theme(
      legend.position = "none",
      panel.grid.major = element_line(colour = "#CCCCCC")
    ) +
    labs(title = paste("Offence locations in", city, "data")) +
    xlim(outline_bbox$xmin - ((outline_bbox$xmax - outline_bbox$xmin) / 5), 
         outline_bbox$xmax + ((outline_bbox$xmax - outline_bbox$xmin) / 5)) +
    ylim(outline_bbox$ymin - ((outline_bbox$ymax - outline_bbox$ymin) / 5), 
         outline_bbox$ymax + ((outline_bbox$ymax - outline_bbox$ymin) / 5))
  cat("\tGenerated plot\n")
  
  ggsave(
    paste0("../temp_data/plot_", 
           str_replace_all(str_to_lower(city), "\\s", "_"), ".pdf"),
    width = 11.69, height = 16.53, units = "in")
  cat("\tSaved plot\n")
  
  rm(city, crimes, outline)
  
}
```

