---
title: "Harmonise data"
output: html_notebook
---

The code in this file harmonises the format of data from each city so that they
can be merged into a single dataset. This is composed of:

 1. renaming variables that should have the same name across all cities
 2. removing redundant variables
 3. changing the type of some variables
 4. adding a prefix to any other variable names while tidying up variable names

The data from each city are not merged in this file because they need to be
checked first. This checking can be done more easily once the data are in a
common format.

The final step needs a list of variables that are not specific to each city and
so do not need a prefix.


# Boston

```{r}
read_rds("../temp_data/spatial_boston_data.Rds") %>% 
  rename(local_row_id = INCIDENT_NUMBER) %>% 
  convert_names(common_vars, cities$prefix[cities$name == "Boston"]) %>% 
  write_rds("../temp_data/final_boston_data.Rds", compress = "gz") %>% 
  glimpse()
```



# Chicago

```{r}
read_rds("../temp_data/spatial_chicago_data.Rds") %>% 
  select(-Date, -`X Coordinate`, -`Y Coordinate`, -Year, -Location) %>% 
  mutate(
    Arrest = as.logical(Arrest),
    Domestic = as.logical(Domestic)
  ) %>% 
  rename(local_row_id = ID) %>% 
  convert_names(common_vars, cities$prefix[cities$name == "Chicago"]) %>% 
  write_rds("../temp_data/final_chicago_data.Rds", compress = "gz") %>% 
  glimpse()
```


# Detroit

```{r}
read_rds("../temp_data/spatial_detroit_data.Rds") %>% 
  select(-INCIDENTDATE, -HOUR) %>% 
  mutate(
    Latitude = ifelse(Latitude < -90 | Latitude > 90, NA, Latitude),
    Longitude = ifelse(Longitude < -180 | Longitude > 180, NA, Longitude),
    CRIMEID = as.numeric(CRIMEID)
  ) %>% 
  rename(local_row_id = CRIMEID, case_number = CRNO) %>% 
  convert_names(common_vars, cities$prefix[cities$name == "Detroit"]) %>% 
  write_rds("../temp_data/final_detroit_data.Rds", compress = "gz") %>% 
  glimpse()
```


# Fort Worth

```{r}
read_rds("../temp_data/spatial_fort_worth_data.Rds") %>% 
  select(-`From Date`, -Offense, -City, -State) %>% 
  rename(
    local_row_id = `Case and Offense`, 
    case_number = `Case Number`, 
    address = `Block Address Expanded`,
    location_type_code = `Location Type`
  ) %>% 
  convert_names(common_vars, cities$prefix[cities$name == "Fort Worth"]) %>% 
  write_rds("../temp_data/final_fort_worth_data.Rds", compress = "gz") %>% 
  glimpse()
```


# Kansas City

```{r}
read_rds("../temp_data/spatial_kansas_city_data.Rds") %>% 
  select(-Reported_Date, -Reported_Time, -From_Date, -From_Time, -To_Date, 
         -To_Time, -City, -Location, -address_string) %>% 
  rename(
    local_row_id = Report_No, 
    NIBRS_Offense_Code = IBRS
  ) %>% 
  mutate_at(vars(one_of("Date.Single", "Date.Start", "Date.End")), format, 
            format = "%Y-%m-%d %H:%M") %>% 
  convert_names(common_vars, cities$prefix[cities$name == "Kansas City"]) %>% 
  write_rds("../temp_data/final_kansas_city_data.Rds", compress = "gz") %>% 
  glimpse()
```


# Los Angeles

```{r}
read_rds("../temp_data/spatial_los_angeles_data.Rds") %>% 
  select(-`Date Occurred`, -`Time Occurred`, -WhichCrime) %>% 
  mutate(Address = str_replace_all(Address, "\\s+", " ")) %>% 
  rename(local_row_id = `DR Number`) %>% 
  convert_names(common_vars, cities$prefix[cities$name == "Los Angeles"]) %>% 
  write_rds("../temp_data/final_los_angeles_data.Rds", compress = "gz") %>% 
  glimpse()
```


# Louisville

```{r}
read_rds("../temp_data/spatial_louisville_data.Rds") %>% 
  select(-DATE_OCCURED, -NIBRS_CODE, -UCR_HIERARCHY, -Offense, -CITY) %>% 
  rename(local_row_id = ID, case_number = INCIDENT_NUMBER, 
         address = BLOCK_ADDRESS) %>% 
  convert_names(common_vars, cities$prefix[cities$name == "Louisville"]) %>% 
  write_rds("../temp_data/final_louisville_data.Rds", compress = "gz") %>% 
  glimpse()
```


# New York

```{r}
read_rds("../temp_data/spatial_new_york_data.Rds") %>% 
  select(-CMPLNT_FR_DT, -CMPLNT_FR_TM, -CMPLNT_TO_DT, -CMPLNT_TO_TM, -X_COORD_CD, 
         -Y_COORD_CD, -Lat_Lon) %>% 
  rename(local_row_id = `CMPLNT_NUM`) %>% 
  convert_names(common_vars, cities$prefix[cities$name == "New York"]) %>% 
  write_rds("../temp_data/final_new_york_data.Rds", compress = "gz") %>% 
  glimpse()
```


# San Francisco

```{r}
read_rds("../temp_data/spatial_san_francisco_data.Rds") %>%
  select(-DayOfWeek, -Date, -Time, -Location) %>% 
  rename(local_row_id = IncidntNum) %>% 
  convert_names(common_vars, cities$prefix[cities$name == "San Francisco"]) %>% 
  write_rds("../temp_data/final_san_francisco_data.Rds", compress = "gz") %>% 
  glimpse()
```


# Seattle

```{r}
read_rds("../temp_data/spatial_seattle_data.Rds") %>% 
  select(-`Occurred Date or Date Range Start`, -`Occurred Date Range End`,
         -`Census Tract 2000`, -Location, -Month, -Year) %>% 
  rename(local_row_id = `RMS CDW ID`, 
         case_number = `General Offense Number`) %>% 
  convert_names(common_vars, cities$prefix[cities$name == "Seattle"]) %>% 
  write_rds("../temp_data/final_seattle_data.Rds", compress = "gz") %>% 
  glimpse()
```


# Tucson

```{r}
read_rds("../temp_data/spatial_tucson_data.Rds") %>% 
  mutate(
    CLEARANCE_DATE = ifelse(is.na(CLEARANCE_DATE), CLEARANCE_, CLEARANCE_DATE),
    UAPremiseDesc = ifelse(is.na(UAPremiseDesc), UAPremiseD, UAPremiseDesc)
  ) %>% 
  select(-OBJECTID, -city, -state, -POINT_X, -POINT_Y, -DATE_OCCU, 
         -HOUR_OCCU, -CLEARANCE_, -UAPremiseD) %>% 
  rename(local_row_id = PrimaryKey, case_number = INCI_ID) %>% 
  convert_names(common_vars, cities$prefix[cities$name == "Tucson"]) %>% 
  write_rds("../temp_data/final_tucson_data.Rds", compress = "gz") %>% 
  glimpse()
```


# Virginia Beach

```{r}
read_rds("../temp_data/spatial_virginia_beach_data.Rds") %>% 
  select(-`Date Occured`, -year) %>% 
  rename(local_row_id = `Police Case Number`) %>% 
  convert_names(common_vars, cities$prefix[cities$name == "Virginia Beach"]) %>% 
  write_rds("../temp_data/final_virginia_beach_data.Rds", compress = "gz") %>% 
  glimpse()
```

