---
title: "Crime Open Database data summary: `r params$city`"
date: "`r format(lubridate::now(), '%e %B %Y')`"
output: pdf_document
always_allow_html: true
header-includes:
   - \usepackage{tabu}
   - \usepackage{booktabs}
   - \usepackage{longtable}
params:
  city: Austin
---

```{r setup, echo = FALSE, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE)

source(here::here("code/01_packages_and_functions.R"))

library("kableExtra")
library("scales")
```

```{r}
this_city <- params$city
this_city_usgs <- recode(
  this_city,
  "Nashville" = "Nashville-Davidson metropolitan government (balance)",
  "St Louis" = "St. Louis"
)

data <- str_glue("temp_data/final_{slugify(this_city, '_')}_data.Rds") %>% 
  here() %>% 
  read_rds()
```

The data for `r this_city` included in the Crime Open Database (CODE)
includes records for `r comma(nrow(data))` offenses between 
`r min(data$date_year)` and `r max(data$date_year)`.



# Excluded cases

```{r}
excluded_cases <- data %>% 
  filter(is.na(latitude) | is.na(longitude) | in_city == FALSE | 
           offense_group == "excluded cases") %>% 
  nrow()
```

Cases are excluded from the final dataset for various reasons. Overall 
`r comma(excluded_cases)` (`r percent((excluded_cases/nrow(data)))` of the 
total) cases have been excluded.

```{r include=TRUE}
data %>% 
  replace_na(list(in_city = FALSE)) %>% 
  mutate(
    Total = TRUE,
    `Incomplete co-ordinates` = is.na(latitude) | is.na(longitude),
    `Location outside city` = !`Incomplete co-ordinates` & !in_city,
    `Non-criminal matters` = !`Incomplete co-ordinates` & 
      !`Location outside city` & offense_code == "99Z",
    `Non-Reportable Crimes` = !`Incomplete co-ordinates` & 
      !`Location outside city` & offense_code == "99Y", 
    `Non-Categorisable Offenses` = !`Incomplete co-ordinates` & 
      !`Location outside city` & offense_code == "99X"
  ) %>% 
  select(Total, `Incomplete co-ordinates`, `Location outside city`, 
         `Non-criminal matters`, `Non-Reportable Crimes`, 
         `Non-Categorisable Offenses`) %>% 
  summarise_all(sum, na.rm = TRUE) %>%
  gather(key = "Reason", value = "Cases") %>%
  replace_na(list(Cases = 0)) %>% 
  mutate(Cases = comma(Cases)) %>% 
  kable(format = "latex", align = "lr", booktabs = TRUE, linesep = "") %>% 
  kable_styling(full_width = TRUE, latex_options = c("repeat_header"))
```

```{r}
## 
## THIS LINE FILTERS THE DATA TO REMOVE EXCLUDED CASES
## If any changes are made above, this line must be re-run before running any
## code below
## 
data <- data %>% 
  filter(
    !is.na(latitude) & 
      !is.na(longitude) & 
      in_city == TRUE & 
      offense_group != "excluded cases"
  )
```

**This leaves a total of `r comma(nrow(data))` genuine cases included in the 
data.**



# Temporal distribution

```{r}
annual_counts <- data %>% 
  select(date_year) %>% 
  group_by(date_year) %>% 
  summarise(`Cases` = n()) %>% 
  rename(Year = date_year)

daily_counts <- data %>% 
  mutate(date = as_date(ymd_hm(date_single))) %>% 
  group_by(date) %>% 
  summarise(n = n())
```

```{r include=TRUE, fig.height=4}
ggplot(daily_counts, aes(x = date, y = n)) +
  geom_line(size = 0.25) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y", 
               expand = expansion(mult = c(0, 0.02))) +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0))) +
  labs(
    title = str_wrap(str_glue(
      "{this_city} had between {min(daily_counts$n)} and {max(daily_counts$n)}",
      " offenses per day (median = {median(daily_counts$n)})"
    )), 
    subtitle = "crimes per day, all offences", 
    x = "date", 
    y = "offenses per day"
  ) +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )
```



```{r include=TRUE, fig.height=4}
data %>% 
  mutate(
    week = week(date_single),
    week = ifelse(week > 52, 1, week)
  ) %>% 
  count(date_year, week) %>% 
  ggplot(aes(x = week, y = n, group = week)) +
  geom_boxplot() +
  # ggthemes::geom_tufteboxplot(median.type = "line", hoffset = 0, size = 0.67) +
  scale_y_continuous(limits = c(0, NA), labels = comma_format()) +
  labs(
    title = "Variation in overall crime frequency through the year", 
    subtitle = "crimes per week, all offences",
    x = "week of the year", 
    y = "offenses per week"
  ) +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )
  
```



```{r include=TRUE, fig.height=4}
daily_counts %>% 
  mutate(weekday = weekdays(date, abbreviate = TRUE)) %>% 
  ggplot(aes(x = weekday, y = n)) +
  geom_violin(draw_quantiles = c(0.50)) +
  ylim(0, NA) +
  labs(
    title = "Variation in overall crime frequency across weekdays", 
    subtitle = "crimes per weekday, all offences", 
    x = "weekday", 
    y = "offenses per day"
  ) +
  scale_x_discrete(limits = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")) +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank()
  )
```

and hours of the day:

```{r include=TRUE, fig.height=4}
data %>% 
  mutate(
    date_time = ymd_hm(date_single),
    date = date(date_time),
    hour = hour(date_time)
  ) %>% 
  count(date, hour) %>% 
  select(-date) %>% 
  ggplot(aes(x = hour, y = n, group = hour)) +
  geom_boxplot() +
  scale_x_continuous(n.breaks = 24, expand = expansion(mult = c(0, 0))) +
  scale_y_continuous(limits = c(0, NA)) +
  labs(
    title = "Variation in overall crime frequency across hours of the day", 
    subtitle = "crimes per hour of the day, all offences", 
    x = "hour", 
    y = "offenses per hour"
  ) +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )
```





# Spatial distribution

```{r message=FALSE, warning=FALSE}
city_outline <- str_glue(
  "spatial_data/cities/tl_2016_{cities$fips[cities$name == this_city]}",
  "_place.shp"
) %>% 
  here() %>% 
  read_sf() %>% 
  filter(NAME == this_city_usgs)

city_bbox <- st_bbox(city_outline)

city_data <- data %>% 
  sample_frac(0.1) %>% 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

city_map <- ggplot() +
  geom_sf(data = tigris::states(year = 2016, class = "sf"), 
          fill = NA, colour = "#999999") +
  geom_hex(
    mapping = aes(x = longitude, y = latitude),
    data = filter(data, !is.na(latitude) & !is.na(longitude)),
    bins = 50,
    na.rm = TRUE
  ) +
  geom_sf(data = city_outline, fill = NA) +
  # geom_sf(data = city_data, size = 0.2, alpha = 0.5) +
  lims(x = c(city_bbox$xmin, city_bbox$xmax), 
       y = c(city_bbox$ymin, city_bbox$ymax)) +
  scale_fill_continuous(low = "#FFFFFF", high = "#000000", 
                        label = scales::comma) +
  labs(title = "Spatial distribution of crimes", 
       subtitle = "all offences, 10% sample", x = "", y = "") +
  theme_minimal() +
  theme(
    panel.grid.major = element_line(colour = "#CCCCCC")
  )
```

```{r include=TRUE, fig.height=8}
city_map
```





# Crime types

Offenses in the CODE data are split into three categories: crimes against people,
property and society. These categories are then split into smaller categories
for different offense types.



```{r include=TRUE, fig.cap="Crimes against people"}
data %>% 
  filter(!is.na(latitude) & !is.na(!longitude) & in_city == TRUE & 
           offense_against == "persons") %>% 
  count(offense_group, offense_code, offense_type, name = "Cases") %>% 
  arrange(offense_code) %>% 
  mutate(
    offense_group = ifelse(
      offense_group != lag(offense_group) | row_number() == 1, 
      offense_group, 
      ""
    ),
    Cases = comma(Cases)
  ) %>% 
  rename(Code = offense_code, Group = offense_group, Type = offense_type) %>% 
  kable(format = "latex", align = "lllr", booktabs = TRUE, longtable = TRUE, 
        linesep = "\\addlinespace", caption = "Crimes against people") %>% 
  kable_styling(full_width = TRUE, latex_options = c("repeat_header")) %>% 
  column_spec(1, width = "5cm") %>% 
  column_spec(2, width = "1cm") %>% 
  column_spec(4, width = "1.5cm")
```



```{r include=TRUE}
data %>% 
  filter(!is.na(latitude) & !is.na(!longitude) & in_city == TRUE & 
           offense_against == "property") %>% 
  count(offense_group, offense_code, offense_type, name = "Cases") %>% 
  arrange(offense_code) %>% 
  mutate(
    offense_group = ifelse(
      offense_group != lag(offense_group) | row_number() == 1, 
      offense_group, 
      ""
    ),
    Cases = comma(Cases)
  ) %>% 
  rename(Code = offense_code, Group = offense_group, Type = offense_type) %>% 
  kable(format = "latex", align = "lllr", booktabs = TRUE, longtable = TRUE, 
        linesep = "\\addlinespace", caption = "Crimes against property") %>% 
  kable_styling(full_width = TRUE, latex_options = c("repeat_header")) %>% 
  column_spec(1, width = "5cm") %>% 
  column_spec(2, width = "1cm") %>% 
  column_spec(4, width = "1.5cm")
```



```{r include=TRUE}
data %>% 
  filter(!is.na(latitude) & !is.na(!longitude) & in_city == TRUE & 
           offense_against == "society") %>% 
  count(offense_group, offense_code, offense_type, name = "Cases") %>% 
  arrange(offense_code) %>% 
  mutate(
    offense_group = ifelse(
      offense_group != lag(offense_group) | row_number() == 1, 
      offense_group, 
      ""
    ),
    Cases = comma(Cases)
  ) %>% 
  rename(Code = offense_code, Group = offense_group, Type = offense_type) %>% 
  kable(format = "latex", align = "lllr", booktabs = TRUE, longtable = TRUE, 
        linesep = "\\addlinespace", caption = "Crimes against society") %>% 
  kable_styling(full_width = TRUE, latex_options = c("repeat_header")) %>% 
  column_spec(1, width = "5cm") %>% 
  column_spec(2, width = "1cm") %>% 
  column_spec(4, width = "1.5cm")
```



```{r include=FALSE}
nibrs_categories <- read_csv(here("crime_categories/nibrs_categories.csv"), 
         col_types = cols(.default = col_character())) %>% 
  filter(
    nibrs_crime_against != "Excluded cases" &
    !nibrs_offense_code %in% c("120", "220")
  )

missing_types <- setdiff(
  nibrs_categories %>% 
    filter(str_sub(nibrs_offense_type, 3, 3) != "U") %>% 
    pluck("nibrs_offense_type") %>% 
    unique() %>% 
    { paste(str_sub(., 0, 3), str_to_lower(str_sub(., 5))) },
  unique(paste(data$offense_code, data$offense_type))
)
```

```{r include=TRUE, results="asis"}
if (length(missing_types) > 0) {
  cat("Some types of crime are missing from the", this_city, "data:\n\n")
  cat(paste("  *", missing_types), sep = "\n")
}
```

```{r include=TRUE, results="asis"}
other_cats <- data %>%
  filter(str_sub(offense_code, 3, 3) == "U") %>%
  mutate(Type = paste(offense_code, offense_type)) %>%
  count(Type, name = "Cases") %>%
  mutate(Cases = comma(Cases))

if (nrow(other_cats) > 0) {
  str_glue(
    "Due to limitations in the matching process between crime types in the ",
    "original data and the CODE crime categories, some crimes have been coded ",
    "to an 'other' category within the relevent higher-level category. ",
    "Specifically, crimes have been coded to the following 'other' ",
    "categories:\n\n"
  )
}
```

```{r include=TRUE}
if (nrow(other_cats) > 0) {
  other_cats %>%
    kable(format = "latex", align = "lr", booktabs = TRUE, 
          linesep = "\\addlinespace", 
          caption = "Crime types missing from the data") %>%
    kable_styling(full_width = TRUE, latex_options = c("repeat_header")) %>%
    column_spec(2, width = "1.5cm")
}
```

