---
title: "City crime data summary"
output:
  html_document:
    code_folding: hide
    df_print: paged
  html_notebook:
    code_folding: hide
  pdf_document: default
---

```{r include=FALSE}
# load libraries (these need to be loaded again because knitr doesn't recognise
# libraries from the global environment)
library('tidyverse')
library("lubridate")
library('sf') # handle spatial data

# for the same reason, load necessary reference objects
cities <- tribble(
  ~name, ~fips, ~prefix,
  "Chicago",        "17", "chi",
  "Detroit",        "26", "dtt",
  "Fort Worth",     "48", "ftw",
  "Los Angeles",    "06", "lax",
  "Louisville",     "21", "lou",
  "New York",       "36", "nyc",
  "San Francisco",  "06", "sfo",
  "Seattle",        "53", "sea",
  "Tucson",         "04", "tus",
  "Virginia Beach", "51", "vib"
)
```


```{r include=FALSE}
this_city <- "Virginia Beach"
data <- read_rds(paste0("../temp_data/final_", 
                        str_replace_all(str_to_lower(this_city), "\\s", "_"), 
                        "_data.Rds"))
```

The data for `r this_city` included in the Crime Open Database (CODE)
includes records for `r format(nrow(data), big.mark = ",")` offences between 
`r min(data$date_year)` and `r max(data$date_year)`.


# Excluded cases

```{r include=FALSE}
excluded_cases <- data %>% 
  filter(is.na(latitude) | is.na(longitude) | in_city == FALSE | 
           offense_group == "excluded cases") %>% 
  nrow()
```

Cases are excluded from the final dataset for various reasons. Overall 
`r format(excluded_cases, big.mark = ",")` 
(`r format((excluded_cases/nrow(data)) * 100, digits = 1, big.mark = ",")`% of 
the total) cases have been excluded.

```{r echo=FALSE}
data %>% mutate(
  Total = TRUE,
  `Incomplete co-ordinates` = 
    ifelse(is.na(latitude) | is.na(longitude), TRUE, FALSE),
  `Location outside city` = 
    ifelse(`Incomplete co-ordinates` == FALSE & in_city == FALSE, TRUE, FALSE),
  `Non-criminal matters` = 
    ifelse(`Incomplete co-ordinates` == FALSE & `Location outside city` == FALSE 
           & offense_code == "99Z", TRUE, FALSE),
  `Non-Reportable Crimes` = 
    ifelse(`Incomplete co-ordinates` == FALSE & `Location outside city` == FALSE 
           & offense_code == "99Y", TRUE, FALSE),
  `Non-Categorisable Offenses` = 
    ifelse(`Incomplete co-ordinates` == FALSE & `Location outside city` == FALSE 
           & offense_code == "99X", TRUE, FALSE)
) %>% 
  select(Total, `Incomplete co-ordinates`, `Location outside city`, 
         `Non-criminal matters`, `Non-Reportable Crimes`, 
         `Non-Categorisable Offenses`) %>% 
  summarise_all(sum) %>%
  gather(key = "Reason", value = "Cases") %>%
  mutate(Cases = format(Cases, big.mark = ","))
```

```{r include=FALSE}
## 
## THIS LINE FILTERS THE DATA TO REMOVE EXCLUDED CASES
## If any changes are made above, this line must be re-run before running any
## code below
## 
data <- data %>% 
  filter(
    !is.na(latitude) & 
      !is.na(longitude) & 
      in_city == TRUE & 
      offense_group != "excluded cases"
  )
```

**This leaves a total of `r format(nrow(data), big.mark = ",")`
geniune cases included in the data.**



# Temporal distribution

```{r include=FALSE}
annual_counts <- data %>% 
  select(date_year) %>% 
  group_by(date_year) %>% 
  summarise(`Cases` = n()) %>% 
  rename(Year = date_year)
```



```{r include=FALSE}
# The mean number of crimes per year between `r min(data$date_year)` and `r max(data$date_year)` was `r format(mean(annual_counts$Cases), digits = 0, big.mark = ",")`, with annual counts:
# `r mutate(annual_counts, Cases = format(Cases, big.mark = ","))`
daily_counts <- data %>% 
  mutate(date = as_date(ymd_hm(date_single))) %>% 
  group_by(date) %>% 
  summarise(n = n())
```

The daily count of crimes varies between `r min(daily_counts$n)` and
`r max(daily_counts$n)` offenses per day, with a median of 
`r median(daily_counts$n)`:

```{r echo=FALSE}
ggplot(daily_counts, aes(x = date, y = n)) +
  geom_line(size = 0.25) +
  xlim(min(data$date_single), NA) +
  ylim(0, NA) +
  labs(title = "Crimes per day", subtitle = "all offences", x = "Date", 
       y = "Offences per day") +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )
```

There is seasonal variation in overall crime:

```{r echo=FALSE}
data %>% 
  mutate(
    week = week(date_single),
    week = ifelse(week > 52, 1, week)
  ) %>% 
  group_by(date_year, week) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  ggplot(aes(x = week, y = n, group = week)) +
  ggthemes::geom_tufteboxplot(median.type = "line", hoffset = 0, size = 0.67) +
  ylim(0, NA) +
  labs(title = "Crimes per week", subtitle = "all offences", 
       x = "Week of the year", y = "Offenses per week") +
  # geom_boxplot(outlier.size = 1, outlier.alpha = 0.5) + 
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )
  
```


There is variation in the frequency of crime across weekdays:

```{r echo=FALSE}
daily_counts %>% 
  mutate(weekday = weekdays(date, abbreviate = TRUE)) %>% 
  ggplot(aes(x = weekday, y = n)) +
  geom_violin(draw_quantiles = c(0.50)) +
  ylim(0, NA) +
  labs(title = "Crimes per weekday", subtitle = "all offences", x = "Weekday", 
       y = "Offenses per day") +
  scale_x_discrete(limits = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")) +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank()
  )
```

and hours of the day:

```{r echo=FALSE}
data %>% 
  mutate(
    date_time = ymd_hm(date_single),
    date = date(date_time),
    hour = hour(date_time)
  ) %>% 
  group_by(date, hour) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  select(-date) %>% 
  # slice(1:200) %>% 
  ggplot(aes(x = hour, y = n, group = hour)) +
  ggthemes::geom_tufteboxplot(median.type = "line", hoffset = 0, size = 0.67) +
  # geom_boxplot(outlier.size = 1, outlier.alpha = 0.5) +
  # ylim(0, 40) +
  ylim(0, NA) +
  labs(title = "Crimes per hour of the day", subtitle = "all offences", 
       x = "Hour", y = "Offenses per hour") +
  # geom_violin(draw_quantiles = c(0.50))
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )
```



# Spatial distribution

```{r fig.height=9, message=FALSE, warning=FALSE, paged.print=FALSE, results="hide"}
city_outline <- st_read(paste0("../spatial_data/cities/tl_2016_", 
                               cities$fips[cities$name == this_city], 
                               "_place.shp"), quiet = TRUE) %>% 
  filter(NAME == this_city)
city_bbox <- st_bbox(city_outline)

ggplot() +
  geom_sf(data = tigris::states(year = 2016, class = "sf"), 
          fill = NA, colour = "#999999") +
  geom_hex(
    mapping = aes(x = longitude, y = latitude),
    data = filter(data, !is.na(latitude) & !is.na(longitude)),
    bins = 50
  ) +
  geom_sf(
    data = city_outline, fill = NA
  ) +
  lims(x = c(city_bbox$xmin, city_bbox$xmax), 
       y = c(city_bbox$ymin, city_bbox$ymax)) +
  scale_fill_continuous(low = "#FFFFFF", high = "#000000", 
                        label = scales::comma) +
  labs(title = "Spatial distribution of crimes", subtitle = "all offences", 
       x = "", y = "") +
  theme_minimal() +
  theme(
    panel.grid.major = element_line(colour = "#CCCCCC")
  )
```


# Crime types

Offenses in the CODE data are split into three categories: crimes against people,
property and society. These categories are then split into smaller categories
for different offense types.

## Crimes against people

```{r echo=FALSE}
data %>% 
  filter(!is.na(latitude) & !is.na(!longitude) & in_city == TRUE & 
           offense_against == "persons") %>% 
  group_by(offense_group, offense_code, offense_type) %>% 
  summarise(Cases = n()) %>% 
  arrange(offense_code) %>% 
  ungroup() %>% 
  select(-offense_code) %>% 
  mutate(
    offense_group = ifelse(nchar(offense_group) > 40, 
                           paste(str_sub(offense_group, 0, 39), "…"), 
                           offense_group),
    offense_type = ifelse(nchar(offense_type) > 40, 
                          paste(str_sub(offense_type, 0, 39), "…"), 
                          offense_type)
  ) %>% 
  mutate(Cases = format(Cases, big.mark = ",")) %>% 
  rename(Group = offense_group, Type = offense_type)
```

## Crimes against property

```{r echo=FALSE}
data %>% 
  filter(!is.na(latitude) & !is.na(!longitude) & in_city == TRUE & 
           offense_against == "property") %>% 
  group_by(offense_group, offense_code, offense_type) %>% 
  summarise(Cases = n()) %>% 
  arrange(offense_code) %>% 
  ungroup() %>% 
  select(-offense_code) %>% 
  mutate(
    offense_group = ifelse(nchar(offense_group) > 40, 
                           paste(str_sub(offense_group, 0, 39), "…"), 
                           offense_group),
    offense_type = ifelse(nchar(offense_type) > 40, 
                          paste(str_sub(offense_type, 0, 39), "…"), 
                          offense_type)
  ) %>% 
  mutate(Cases = format(Cases, big.mark = ",")) %>% 
  rename(Group = offense_group, Type = offense_type)
```

## Crimes against society

```{r echo=FALSE}
data %>% 
  filter(!is.na(latitude) & !is.na(!longitude) & in_city == TRUE & 
           offense_against == "society") %>% 
  group_by(offense_group, offense_code, offense_type) %>% 
  summarise(Cases = n()) %>% 
  arrange(offense_code) %>% 
  ungroup() %>% 
  select(-offense_code) %>% 
  mutate(
    offense_group = ifelse(nchar(offense_group) > 40, 
                           paste(str_sub(offense_group, 0, 39), "…"), 
                           offense_group),
    offense_type = ifelse(nchar(offense_type) > 40, 
                          paste(str_sub(offense_type, 0, 39), "…"), 
                          offense_type)
  ) %>% 
  mutate(Cases = format(Cases, big.mark = ",")) %>% 
  rename(Group = offense_group, Type = offense_type)
```

## Crime types missing from the data

```{r include=FALSE}
nibrs_categories <- read_csv("../crime_categories/NIBRS categories.csv", 
         col_types = cols(.default = col_character())) %>% 
  filter(
    NIBRS_Crime_Against != "Excluded cases" &
    !NIBRS_Offense_Code %in% c("120", "220")
  ) %>% 
  select(-NIBRS_Offense_Code_1)

missing_types <- setdiff(
  nibrs_categories %>% 
    filter(str_sub(NIBRS_Offense_Type, 3, 3) != "U") %>% 
    { unique(str_to_title(.$NIBRS_Offense_Type)) },
  unique(str_to_title(paste(data$offense_code, data$offense_type)))
)
```

```{r results="asis"}
if (length(missing_types) > 0) {
  cat("Some types of crime are missing from the", this_city, "data:\n\n")
  cat(paste("  *", missing_types), sep = "\n")
}
```

```{r results="asis"}
other_cats <- data %>% 
  filter(str_sub(offense_code, 3, 3) == "U") %>% 
  mutate(Type = paste(offense_code, offense_type)) %>% 
  group_by(Type) %>% 
  summarise(Cases = n()) %>% 
  mutate(Cases = format(Cases, big.mark = ","))

if (nrow(other_cats) > 0) {
  cat("Due to limitations in the matching process between crime types in the",
      "original data and the CODE crime categories, some crimes have been", 
      "coded to an 'other' category within the relevent higher-level category.", 
      "Specifically, crimes have been coded to the following 'other'", 
      "categories:", "\n\n")
}
```

```{r echo=FALSE}
if (nrow(other_cats) > 0) {
  other_cats
}
```

