---
title: "Merge and output data"
output: html_notebook
---


# Output RDS files

This code outputs RDS files for each city for each year. These can be used to
produce CSV files for multiple cities etc.

Three files are produced for each city for each year. The 'extended' file 
contains all the variables, the 'core' file contains only the common variables 
(i.e. those without the city-name prefix) and the 'sample' file contains a 1% 
sample of data in the core file.

```{r}
# remove unnecessary variables from common_vars
common_vars <- setdiff(
  common_vars, 
  c('geometry', "multiple_dates", "in_city", "address", "fips_state", 
    "fips_county", "tract", "block_group", "block", "date_year")
) %>% 
  # rename the crime category variables because they've been renamed in the 
  # dataset by the convert_names() function in file 05_harmonise_data.Rmd
  mapvalues(c('nibrs_offense_code', 'nibrs_offense_type', 
              'nibrs_offense_category', 'nibrs_crime_against'), 
            c('offense_code', 'offense_type', 'offense_group', 
              'offense_against')) %>% 
  # add census_block to common vars so it appears in the core data
  append("census_block")
```

```{r}
# set row number, which should be continuous across cities and years so must be
# remembered between files
row_num <- 1

# for each city, load the data and output an Rds file for each year
# this must be done separately for each city to avoid the 'vector memory
# exhausted' problem on macOS described at
# http://r.789695.n4.nabble.com/R-3-5-0-vector-memory-exhausted-error-on-readBin-td4750237.html
for (city in cities$name) {
  
  # load and process data
  city_data <- readRDS(paste0("../temp_data/final_", 
                              str_replace_all(str_to_lower(city), "\\s", "_"), 
                              "_data.Rds")) %>% 
    mutate_at(vars(one_of(c("local_row_id", "case_number"))), as.character) %>% 
    arrange(date_single) %>%
    filter(
      offense_against != "excluded cases" 
      & !is.na(longitude) 
      & !is.na(latitude) 
      & in_city == TRUE
    ) %>% 
    mutate(
      # add a row ID
      uid = row_num:(row_num + nrow(.) - 1),
      # merge the census identifiers
      census_block = paste0(fips_state, fips_county, tract, block),
      # catch and robbery/burglary offenses have have not been recategorised
      offense_code = case_when(
        offense_code == "120" ~ "12U",
        offense_code == "220" ~ "22U",
        offense_code == "23U" ~ "23H",
        TRUE ~ offense_code
      )
    ) %>% 
    select(one_of(common_vars), everything(), -in_city, -multiple_dates,
           -fips_state, -fips_county, -tract, -block_group, -block) %>% 
    mutate(year = year(date_single))
  
  # output data by year
  city_data %>% 
    split(.$year) %>% 
    walk(function (x) {

      this_year <- x$year[1]
      this_city <- str_replace_all(str_to_lower(city), "\\s", "_")

      x %>%
        select(-year) %>% 
        # export extended data - use write_rds() rather than saveRDS() because
        # it returns the input invisibly for use inside a pipe
        write_rds(path = paste0("../output_data/crime_open_database_extended_",
                                this_city, "_", this_year, ".Rds"),
                  compress = "bz2") %>%
        # select core variables and export
        select(one_of(common_vars)) %>%
        write_rds(path = paste0("../output_data/crime_open_database_core_",
                                this_city, "_", this_year, ".Rds"),
                  compress = "bz2") %>% 
        # take a 1% sample and export
        sample_frac(0.01) %>% 
        arrange(uid) %>% 
        write_rds(path = paste0("../output_data/crime_open_database_sample_",
                                this_city, "_", this_year, ".Rds"),
                  compress = "bz2")

      message("\nWritten data for ", city, " in ", this_year)
      
    })
  
  row_num <- row_num + nrow(city_data)
  
}

# clean up
rm(city, city_data, row_num)
```


## Check that UIDs are consecutive across files

```{r}
dir("../output_data/", 
                 pattern = "^crime_open_database_core_(.+?).Rds$", 
                 full.names = TRUE) %>% 
  map_df(function (x) {
    read_rds(x) %>% 
      summarise(city = first(city_name), year = year(first(date_single)), 
                uid_min = min(uid), uid_max = max(uid))
  }) %>% 
  mutate(uid_ok = ifelse(uid_min - 1 == lag(uid_max), "OK", "not OK")) %>% 
  filter(complete.cases(.)) %>% 
  group_by(city, uid_ok) %>% 
  summarise(n = n()) %>% 
  spread(uid_ok, n)
```


# Merge data by year

RDS files can only be used by R, so gzipped CSV files are also needed for those
using other software. Since most people will be downloading these manually, the
data from all cities are aggregated into one file per year.

```{r}
for (file_type in c("core", "extended", "sample")) {
  
dir("../output_data", 
    pattern = paste0("^crime_open_database_", file_type, "_(.+?).Rds$"), 
    full.names = TRUE) %>% 
  str_extract("\\d+") %>% 
  unique() %>% 
  walk(function (x) {
    
    dir("../output_data", 
        pattern = paste0("^crime_open_database_", file_type, "_(.+?)", x, 
                         ".Rds$"), 
        full.names = TRUE) %>% 
      map_df(read_rds) %>% 
      write_rds(path = paste0("../output_data/crime_open_database_", file_type, 
                              "_all_", x, ".Rds"),
                compress = "bz2") %>% 
      write.csv(
        gzfile(paste0("../output_data/crime_open_database_", file_type, "_", x,
                      ".csv.gz")),
        na = "", row.names = FALSE, fileEncoding = "UTF-8")
    
    message("Merged ", file_type, " data for ", x, " into one RDS and one CSV ",
            "file")

  })
  
}

rm(file_type)  
```


# Report typical file sizes

```{r}
cat(
  "\nMean size for core-data files:", 
  dir("../output_data", pattern = "^crime_open_database_core_(.+?).csv.gz$", 
      full.names = TRUE) %>% 
    file.size() %>% 
    { round(mean(.) / 10^6, 1) },
  "MB\nMean size for extended-data files:",
  dir("../output_data", pattern = "^crime_open_database_extended_(.+?).csv.gz$", 
      full.names = TRUE) %>% 
    file.size() %>% 
    { round(mean(.) / 10^6, 1) },
  "MB\nMean size for sample-data files:",
  dir("../output_data", pattern = "^crime_open_database_sample_(.+?).csv.gz$", 
      full.names = TRUE) %>% 
    file.size() %>% 
    { round(mean(.) / 10^6, 1) },
  "MB\n"
)
```

