---
title: "Merge and output data"
output: html_notebook
---


# Merge data

This code merges the data from each city into one object, which can then be used
to output the final data.

```{r}
# rename the crime category variables because they've been renamed in the 
# dataset by the convert_names() function in file 05_harmonise_data.Rmd
common_vars <- setdiff(common_vars, c('geometry')) %>% 
  mapvalues(c('nibrs_offense_code', 'nibrs_offense_type', 
              'nibrs_offense_category', 'nibrs_crime_against'), 
            c('offense_code', 'offense_type', 'offense_group', 
              'offense_against'))

# load and merge the dataset for each city
data <- dir("../temp_data", pattern = "^final_", full.names = TRUE) %>% 
  # report_status(paste("Identified files in directory at", 
  #                     as.character(now))) %>% 
  map(read_rds) %>% 
  # report_status(paste("Read data from files at", as.character(now()))) %>% 
  map(mutate_at, vars(one_of(c("local_row_id", "case_number"))), 
      as.character) %>% 
  # map(mutate, 
  #     local_row_id = as.character(local_row_id),
  #     case_number = as.character(case_number)) %>% 
  # report_status(paste("Changed identifiers to character vectors at", 
  #                     as.character(now()))) %>% 
  bind_rows() %>% 
  # report_status(paste("Merged data from individual files at", 
  #                     as.character(now()))) %>% 
  arrange(city_name, date_single) %>% 
  # report_status(paste("Arranged data by city and date at", 
  #                     as.character(now()))) %>% 
  filter(
    offense_against != "excluded cases" & 
    !is.na(longitude) &
    !is.na(latitude) &
    in_city == TRUE
  ) %>%
  # report_status(paste("Removed excluded cases at", as.character(now()))) %>% 
  mutate(
    uid = 1:nrow(.)
  ) %>% 
  mutate(
    offense_code = case_when(
      offense_code == "120" ~ "12U",
      offense_code == "220" ~ "22U",
      offense_code == "23U" ~ "23H",
      TRUE ~ offense_code
    )
  ) %>% 
  # report_status(paste("Added unique row ID at", as.character(now()))) %>% 
  select(one_of(common_vars), everything(), -in_city, -multiple_dates)

# data %>%
#   sample_n(50) %>%
#   arrange(city_name, date_single) %>%
#   View()

cat("CODE contains data on", format(nrow(data), big.mark = ","), "offenses\n")
```

Overall, there are `r nrow(data)` offences in the dataset. This can be broken
down by city:

```{r}
data %>% 
  group_by(city_name, date_year) %>% 
  summarise(n = n()) %>% 
  summarise(crimes_per_year = format(mean(n), scientific = FALSE, digits = 0, 
                                     big.mark = ","))
```

and by city for each year:

```{r}
data %>% 
  group_by(city_name, date_year) %>% 
  summarise(n = n()) %>% 
  spread(date_year, n)
```


# Output data

For each year, two files should be exported: one containing all the variables 
and one containing only the common variables (i.e. those without the city-name 
prefix).

```{r}
# If yearFirst has been set to a very early year, there may be no data for the 
# first few years, producing empty CSV files. To stop this, reset yearFirst if 
# it is before the first year present in the dataset.
if (min(data$date_year) > yearFirst) {
  yearFirst <- min(data$date_year)
}

for (i in yearFirst:yearLast) {
  for (j in c('extended', 'core')) {
    if (j == 'core') {
      these_vars = common_vars
    } else {
      these_vars = names(data)
    }
    file_name <- paste0("../output_data/crime_open_database_", j, "_", i, 
                        ".csv.gz")
    data %>%
      filter(date_year == i) %>%
      select_if(names(data) %in% these_vars) %>%
      select(-date_year) %>% 
      write.csv(gzfile(file_name), na = "", row.names = FALSE)
    cat("Saved file", paste0("'", file_name, "',"), 
        round(file.size(file_name) / 10^6), "MB\n")
  }
}

for (i in yearFirst:yearLast) {
  file_name <- paste0("../output_data/crime_open_database_sample_", i, 
                      ".csv.gz")
  data %>% 
    filter(date_year == i) %>% 
    select_if(names(data) %in% common_vars) %>% 
    select(-date_year) %>% 
    sample_frac(0.01) %>% 
    write.csv(gzfile(file_name), na = "", row.names = FALSE, 
              fileEncoding = "UTF-8")
  cat("Saved file", paste0("'", file_name, "',"), 
      round(file.size(file_name) / 10^6), "MB\n")
}

# clean up
rm(common_vars, these_vars, i, j)

```


## Create sample data

Since the data files are so large, 1% samples may be useful for exploratory
analysis.

```{r}
dir("../output_data", 
    pattern = "^crime_open_database_core", full.names = TRUE) %>%
  purrr::map(function (x) {
    
    new_file_name <- stringr::str_replace(x, "core", "sample")
    
    readr::read_csv(x, col_types = readr::cols(
      .default = readr::col_character(),
      uid = readr::col_integer(),
      date_single = readr::col_datetime(format = ""),
      longitude = readr::col_double(),
      latitude = readr::col_double(),
      fips_state = readr::col_integer(),
      block_group = readr::col_integer(),
      block = readr::col_integer()
    )) %>%
      dplyr::sample_frac(0.01) %>%
    readr::write_csv(new_file_name, na = "")
    
    cat("Saved file", paste0("'", new_file_name, "',"), 
        round(file.size(new_file_name) / 10^6), "MB\n")
    
  })
```


## Report typical file sizes

```{r}
# report typical file sizes
cat(
  "\nMean size for core-data files:", 
  dir("../output_data", pattern = "^crime_open_database_core_(.+?).csv.gz$", 
      full.names = TRUE) %>% 
    file.size() %>% 
    { round(mean(.) / 10^6) },
  "MB\nMean size for extended-data files:",
  dir("../output_data", pattern = "^crime_open_database_extended_(.+?).csv.gz$", 
      full.names = TRUE) %>% 
    file.size() %>% 
    { round(mean(.) / 10^6) },
  "MB\nMean size for sample-data files:",
  dir("../output_data", pattern = "^crime_open_database_sample_(.+?).csv.gz$", 
      full.names = TRUE) %>% 
    file.size() %>% 
    { round(mean(.) / 10^6) },
  "MB\n"
)
```


## Produce Rds files

Rds files are used for downloading within the crimedata R package because they
are smaller, albeit not interoperable with other software.

```{r}
dir("../output_data", 
    pattern = ".csv.gz$", full.names = TRUE) %>%
  purrr::map(function (x) {
    
    new_file_name <- stringr::str_replace(x, ".csv.gz", ".Rds")

    readr::read_csv(x, col_types = readr::cols(
      .default = readr::col_character(),
      uid = readr::col_integer(),
      date_single = readr::col_datetime(format = ""),
      longitude = readr::col_double(),
      latitude = readr::col_double(),
      fips_state = readr::col_integer(),
      block_group = readr::col_integer(),
      block = readr::col_integer()
    )) %>%
      dplyr::select(-local_row_id, -case_number) %>% 
      saveRDS(new_file_name, compress = "bzip2")
    
    cat("Saved file", paste0("'", new_file_name, "',"), 
        round(file.size(new_file_name) / 10^6), "MB\n")
    
  })
```


